---
title: "Introduction to collocation functions in the quanteda.extras R package"
author: "David Brown"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{collocations_introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

## Load the quanteda.extras package

Load the package, as well as others that we'll use in this vignette.

```{r setup, message = FALSE, error=FALSE, warning=FALSE}
library(quanteda.extras)
library(quanteda)
library(tidyverse)
library(ggraph)
```

## Prepare the data

First, we'll use to **preprocess_text( )** function to "clean" the text data. See the **preprocess vignette** for more information.

```{r data_prep, message = FALSE, error=FALSE}
sc <- sample_corpus %>%
  mutate(text = preprocess_text(text))
```

Next, create a corpus and tokenize the data:

```{r tokens}
sc_tokens <- sc %>%
  corpus() %>% # create a corpus object
  tokens(what="fastestword", remove_numbers=TRUE) # tokenize the data
```

## Collocates by mutual information

The **collocates_by_MI( )** function produces collocation measures (by pointwise mutual information) for a specified token in a **quanteda tokens** object. In addition to a token, a span or window (as given by a number of words to the **left** and **right** of the **node word**) is required. The default is 5 to the left and 5 to the right.

```{r coll_money}
money_collocations <- collocates_by_MI(sc_tokens, "money")
```

Check the result:

```{r echo=FALSE}
knitr::kable(head(money_collocations))
```


```{r}
time_collocations <- collocates_by_MI(brown_tkns, "time")

tc <- time_collocations %>% filter(col_freq >= 5 & MI_1 >= 5)
mc <- money_collocations %>% filter(col_freq >= 5 & MI_1 >= 5)
```


```{r}
net <- col_network(tc, mc)
```

```{r}
ggraph(net, weight = link_weight, layout = "stress") + 
  geom_edge_link(color = "gray80", alpha = .75) + 
  geom_node_point(aes(alpha = node_weight, size = 3, color = n_intersects)) +
  geom_node_text(aes(label = label), repel = T) +
  scale_alpha(range = c(0.2, 0.9)) +
  theme_graph() +
  theme(legend.position="none")
```

