---
title: "Introduction to collocation functions in the quanteda.extras R package"
author: "David Brown"
output: rmarkdown::html_vignette
bibliography: collocation_bib.bib
link-citations: yes
nocite: |
  @*
vignette: >
  %\VignetteIndexEntry{collocations_introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

## Load the quanteda.extras package

Load the package, as well as others that we'll use in this vignette.

```{r setup, message = FALSE, error=FALSE, warning=FALSE}
library(quanteda.extras)
library(quanteda)
library(tidyverse)
library(ggraph)
```

## Prepare the data

First, we'll use to **preprocess_text( )** function to "clean" the text data. See the **preprocess vignette** for more information.

```{r data_prep, message = FALSE, error=FALSE}
sc <- sample_corpus %>%
  mutate(text = preprocess_text(text))
```

Next, create a corpus and tokenize the data:

```{r tokens}
sc_tokens <- sc %>%
  corpus() %>% # create a corpus object
  tokens(what="fastestword", remove_numbers=TRUE) # tokenize the data
```

## Collocates by mutual information

The **collocates_by_MI( )** function produces collocation measures (by pointwise mutual information) for a specified token in a **quanteda tokens** object. In addition to a token, a span or window (as given by a number of words to the **left** and **right** of the **node word**) is required. The default is 5 to the left and 5 to the right.

We'll start by making a table of tokens that collocate with the token *money*.

```{r coll_money}
money_collocations <- collocates_by_MI(sc_tokens, "money")
```

Check the result:

```{r echo=FALSE}
knitr::kable(head(money_collocations))
```

Now, let's make a similar table for collocates of *time*.

```{r coll_time}
time_collocations <- collocates_by_MI(sc_tokens, "time")
```


```{r echo=FALSE}
knitr::kable(head(time_collocations))
```

As is clear from the above table, MI is sensitive to rare/infrequent words. Because of that sensitivity, it commmon to make thresholds for both token frequency (absolute frequency) and MI score (ususally at some value ≥ 3).

For our purposes, we'll filter for AF ≥ 5 and MI ≥ 5.

```{r filter}
tc <- time_collocations %>% filter(col_freq >= 5 & PMI >= 5)
mc <- money_collocations %>% filter(col_freq >= 5 & PMI >= 5)
```

Check the result:

```{r echo=FALSE}
knitr::kable(head(tc))
```

```{r echo=FALSE}
knitr::kable(head(mc))
```

## Create a tbl_graph object for plotting

A [**tbl_graph**](https://www.data-imaginist.com/2017/introducing-tidygraph/) is a data structure for **tidyvers** (ggplot2) network plotting.

For this, we'll use the **col_network( )** function.

```{r net}
net <- col_network(tc, mc)
```


## Plot network

The network plot shows the tokens that distinctly collocate with either *time* or *money*, as well as those that interect. The distance from the central tokens (*time* and *money*) is governed by the MI score and the transparency (or alpha) is governed by the token frequency.

The aesthetic details of the plot can be manipulated in the various **ggraph** options.

```{r net_plot, message = FALSE, fig.width = 7, fig.height=4}
ggraph(net, weight = link_weight, layout = "stress") + 
  geom_edge_link(color = "gray80", alpha = .75) + 
  geom_node_point(aes(alpha = node_weight, size = 3, color = n_intersects)) +
  geom_node_text(aes(label = label), repel = T, size = 3) +
  scale_alpha(range = c(0.2, 0.9)) +
  theme_graph() +
  theme(legend.position="none")
```

## Bibliography


