---
title: "Introduction to dispersion functions in the quanteda.extras R package"
author: "David Brown"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dispersions_introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

## Load the quanteda.extras package

Load the package, as well as others that we'll use in this vignette.

```{r setup, message = FALSE, error=FALSE, warning=FALSE}
library(quanteda.extras)
library(quanteda)
library(tidyverse)
```

## Prepare the data

First, we'll use to **preprocess_text( )** function to "clean" the text data. See the **preprocess vignette** for more information.

```{r data_prep, message = FALSE, error=FALSE}
sc <- sample_corpus %>%
  mutate(text = preprocess_text(text))
```

Next, create a corpus and tokenize the data:

```{r tokens}
sc_tokens <- sc %>%
  corpus() %>% # create a corpus object
  tokens(what="fastestword", remove_numbers=TRUE) # tokenize the data
```

## Frequency table

The **frequency_table( )** function aggregates useful descriptive measures: absolute frequency, relative frequency, average reduced frequency, and deviation of proportions.

The relatvie frequency (**Per_10.x**) the relatvie frequency is automatically calibrated to a normaizing factor, where here is per million tokens.

Average reduced frequency (**ARF**) combines dispersion and frequency into a single measure by de-emphasizing occurrences of a token that appear clustered in close proximity.

[Deviation of proportions](https://www.researchgate.net/publication/233685362_Dispersions_and_adjusted_frequencies_in_corpora) (**DP**) is a dispersion statistic proposed by Greis, which measures dispersion on a scale of 0 to 1 such that tokens with DP close to zero are **more** dispersed while those closer to 1 are **less** dispersed.

```{r freq_table, message = FALSE, error=FALSE, warning=FALSE}
ft <- frequency_table(sc_tokens)
```

Check the result:

```{r echo=FALSE}
knitr::kable(head(ft))
```

## Dispersion statstics for all tokens

A data.frame of common dispersion measures can be generated using the **dispersions_all( )** function. The table includes:

* Carroll's *D*<sub>2</sub>
* Rosengren's *S*
* Lyne's *D*<sub>3</sub>
* Distributional Consistency (DC)
* Juilland' *D*
* Deviation of Proportions
* Deviation of Proportions *Norm*

To create the table, the the **quanteda tokens** must first be converted into a **dfm**:

```{r disp_table, message = FALSE, error=FALSE, warning=FALSE}
sc_dfm <- sc_tokens %>% dfm()
dt <- dispersions_all(sc_dfm)
```

Check the result:

```{r echo=FALSE}
knitr::kable(head(dt))
```

## Dispersion statistics for a singe token

The **dispersions_token( )** function calculates the dispersion measures for a single token. It returns a named list with all of the dispersion measures discussed by [S.T. Gries](http://www.stgries.info/research/dispersion/links.html) and the function is baesed on a script he originally authored.

```{r disp_token}
dispersions_token(sc_dfm, "data") %>%
  unlist() %>% 
  knitr::kable(col.names = "")
```


