---
title: "Introduction to pre-processing functions in the quanteda.extras R package"
author: "David Brown"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{preprocess_introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```


```{r setup, message = FALSE, error=FALSE, warning=FALSE}
library(quanteda.extras)
library(tidyverse)
```


## Pre-processing

The  **preprocess_text( )** function takes the following logical (TRUE/FALSE) arguments:

* **contractions** (if set to TRUE contractions will be separated so that, for example, *can't* becomes *ca n't*)
* **hyphens** (if set to TRUE hyphens will be replaced by spaces)
* **punctuation** (if set to TRUE all punctuation marks will be exluded)
* **lower_case** (if set to TRUE all strings are converted to lower case)
* **accent_replace** (if set to TRUE accented chacaracters will be replaced by unaccented ones)
* **remove_numers** (if set to TRUE strings made up of numbers will be eliminated)

### contractions:

```{r contractions}
a <- preprocess_text("can't won't we'll its' it's")
b <- preprocess_text("can't won't we'll its' it's", contractions = FALSE)
```

```{r echo=FALSE}
knitr::kable(a, col.names = "TRUE")
knitr::kable(b, col.names = "FALSE")
```

### hyphens:

```{r}
a <- preprocess_text("un-knowable bluish-gray slo-mo stop-")
b <- preprocess_text("un-knowable bluish-gray slo-mo stop-", hypens = FALSE)
```

```{r echo=FALSE}
knitr::kable(a, col.names = "TRUE")
knitr::kable(b, col.names = "FALSE")
```

### punctuation:

```{r}
a <- preprocess_text("u.k. 50% 'cat' #great now?")
b <- preprocess_text("u.k. 50% 'cat' #great now?", punctuation = FALSE)
```

```{r echo=FALSE}
knitr::kable(a, col.names = "TRUE")
knitr::kable(b, col.names = "FALSE")
```

### lower_case:

```{r}
a <- preprocess_text("U.K. This A-1 1-A")
b <- preprocess_text("U.K. This A-1 1-A", lower_case = FALSE)
```

```{r echo=FALSE}
knitr::kable(a, col.names = "TRUE")
knitr::kable(b, col.names = "FALSE")
```

### accent_replace:

```{r}
a <- preprocess_text("fiancée naïve façade")
b <- preprocess_text("fiancée naïve façade", accent_replace = FALSE)
```

```{r echo=FALSE}
knitr::kable(a, col.names = "TRUE")
knitr::kable(b, col.names = "FALSE")
```

### remove_numbers:

```{r}
a <- preprocess_text("a-1 b2 50% 99 10,000", remove_numbers = TRUE)
b <- preprocess_text("a-1 50% 99 10,000")
```

```{r echo=FALSE}
knitr::kable(a, col.names = "TRUE")
knitr::kable(b, col.names = "FALSE")
```

Note that these options represent some procedures that are common when "cleaning" texts. They give additional control over how a corpus is later "tokenized". These are *not* intended to be comprehensive. Depending on one's data there may be other, specific ways a corpus needs to be processed prior to tokenizing.

The [**textclean**](https://github.com/trinker/textclean) package offers a host of options for pre-processing tasks. In addition, the [**tokens( )**](http://quanteda.io/reference/tokens.html) function in **quanteda** has a variety of built-in options, some similar to the ones described above.

And, of course, one can use either native R **gsub( )** or [**stringr**](https://stringr.tidyverse.org/) **tidyverse** to create task-specific text processing functions.

## Sample corpus

The package comes with a small corpus -- the **sample_corpus**. The corpus contains data from 8 text-types:

* Academic
* Blog
* Fiction
* Magazine
* News
* Spoken
* Television & Movie
* Web

In this way, it resembles the [Corpus of Contemporary American English](https://www.english-corpora.org/coca/). However, it contains only 50 texts from each type, and each text is only about 2,500 words. Thus, it is similar to the Brown family of corpora in its size (roughly 1 million words).

Note that this data is included *only* for demonstration purposes. It was *not* compiled to be used for research.

```{r}
sc <- sample_corpus %>%
  mutate(text_type = str_extract(doc_id, "[a-z]+")) %>%
  group_by(text_type) %>%
  tally()
```


```{r echo=FALSE}
knitr::kable(sc)
```

